Optimal updating factor in Wang-Landau and metadynamics simulations

### 2016-01-25 doc/invt.tex
adding a few references

The Wang-Landau (WL) algorithm\cite{wang2001, wang2001pre}
and metadynamics\cite{laio2002}
are two widely-used techniques for this purpose,
in which a bias potential is actively built-up on-the-fly
during simulation.
%
In essence, these techniques
encourage the system to leave a region that has been visited
by elevating the bias potential there.
%
A main difference of the two techniques
is that in the WL case, an update is limited to the bin
containing the current $z$,
while metadynamics adopts a more widespread window
covering several neighboring bins.
%
The latter ensures a smooth profile
suitable for numerical differentiation for force calculation,
and hence is more often used molecular dynamics simulations.
%
To be effective, the updates to bias potential
are made at every few steps,
while a side effect is that the underlying dynamics
is no longer a equilibrium sampling
required by statistical mechanics.
%
Thus, one has to reduce the magnitude of updating
over the course of simulation.

Naturally, the manner of reducing
the updating magnitude\cite{
belardinelli2007, belardinelli2007jcp, belardinelli2008,
morozov2007, zhou2008, morozov2009,
komura2012, caparica2012, caparica2014}
determines the precision of the final bias potential,
hence that of PMF.
%
For the WL algorithm, it is found
that asymptotically the optimal magnitude of updating $\ln f$
should be given by the
inverse time\cite{
belardinelli2007, belardinelli2007jcp, belardinelli2008,
morozov2007, zhou2008},
where ``time'' is defined as
the number of simulation steps
divided by the number of bins.

@@

Generally, we may wish to sample
a nearly flat distribution\cite{dayal2004, trebst2004}, $p_i$,
by adjusting the bias potential $V_i$
on the fly.
%
For example,
we can monitor the histogram accumulated over
a period of time and use it for $\pi_i$ in
Eq. \eqref{eq:pi_p_phi1},
and then update
$V_i$ according to
$$
V_i^{\mathrm{new}}
=
V_i^{\mathrm{old}}
-
\log \frac{ p_i } { \pi_i }.
$$
%
%From Eq. \eqref{eq:pi_p_phi1},
%it is clear that to achieve the target distribution
%$\pi_i = p_i$ requires
%%
%\begin{equation}
%\log p^*_i
%=V_i + \ln p_i + c
%\label{eq:bias_pmf}
%\end{equation}
%%
%%
%In the end of the simulation,
%the PMF, $-\log p^*_i$,
%can be deduced from Eq. \eqref{eq:bias_pmf}.
%
The above inversion process,
adopted by early multicanonical or
entropic sampling\cite{berg1992, lee1993},
can be inconvenient in practice.
%
For example, a reasonably precise the histogram
requires an accumulation period
longer than the autocorrelation time,
whereas the latter is difficult to
estimate beforehand for an unknown system.

@@

In the WL algorithm\cite{wang2001, wang2001pre},
$v_i$ are updated
in each MC step $t$,
and $v_i(t)$ as a function of $t$
is updated as
%
\begin{equation}
  v_i(t+1)
  =
  v_i(t)
  +
  \delta_{i, \, i(t)}
  \frac{ \alpha(t) } { p_i }.
  \label{eq:wl_update}
\end{equation}
%
where $i(t)$ is the bin at step $t$,
and $\alpha(t)$ is the updating magnitude.
%
We refer to the above scheme as the \emph{single-bin scheme},
for it applies only to the bias potential
at the current bin, $i(t)$,
%
In a more general \emph{multiple-bin scheme}, however,
several neighboring bins are updated:
%
\begin{equation}
  v_i(t+1)
  =
  v_i(t)
  +
  w_{i, \, i(t)}
  \frac{ \alpha(t) } { p_i },
  \label{eq:mbin_update}
\end{equation}
%
where $w_{ij}$ are a set of coefficients satisfying
$\sum_{j=1}^n w_{ij} = 1$.
%
We shall discuss other restrictions on $w_{ij}$
in Sec. \ref{sec:updating-matrix}.

One can show that with a constant $\alpha(t) = \alpha > 0$,
the distribution collected from
the trajectory is identical to $p_i$.
%
However, the above updates
change $v_i(t)$ continuously,
and Eq. \eqref{eq:pi_p_v} no longer holds
at all times.
%%
%The source of the deviation is two-fold.
%%
%First, since $v_i(t)$ is updated continuously,
%there is a random noise that is proportional
%to $\sqrt \alpha$.
%%
%Second, there is a systematic error
%that comes from the updating dynamics itself,
%as it breaks the Markovian nature
%that underlies Eq. \eqref{eq:pi_p_v}.
%%
%
Thus, one has to decrease $\alpha$ over time
to reduce the error of $v_i$,
in order to get the unbiased distribution
$\ln p_i^*$ via Eq. \eqref{eq:v_def}.

In the original WL scheme,
the updating magnitude $\alpha$ (or $\ln f$
in the original paper) is kept as a constant
for a period of time,
which is referred to as a stage below.
%
The histogram collected in the stage is monitored.
%
Once the histogram is sufficiently flat,
we are allowed to enter a new stage
with a reduced $\alpha$\cite{wang2001, wang2001pre}
(usually half as large as
that in the previous stage).
%
This scheme works well for early stages.
%
However, in later stages, it tends to reduce $\alpha$
too quickly, making the asymptotic error
saturate\cite{belardinelli2007, belardinelli2007jcp, belardinelli2008}.

@@

Particularly, if $\mu_2 = \dots = \mu_b = 0$,
the matrix $\mathbf w$ is tridiagonal,
%
\begin{equation}
\arraycolsep=3.6pt\def\arraystretch{1.4}
\mathbf w
=
\left(
  \begin{array}{cccccccc}
    1 - \mu_1   & \mu_1 & 0 & \dots & 0 \\
    \mu_1 & 1 - 2 \, \mu_1  & \mu_1 & \dots & 0 \\
    \vdots & &  & & \vdots \\
    0 & \dots & \mu_1 & 1 - 2 \, \mu_1  & \mu_1 \\
    0 & \dots & 0 & \mu_1 & 1 - \mu_1
  \end{array}
\right).
\label{eq:wtridiag}
\end{equation}
%
We have
\begin{equation}
  \lambda_k = 1 - 4 \, \mu_1 \sin^2 \frac{(k - 1) \, \pi}{2n}.
\label{eq:wtridiag_eigenvalue}
\end{equation}
%
Thus, for this updating matrix to be stable,
we need
$\min \lambda_k = \lambda_n > 0$,
and
$$
\mu_1 < \frac{1}{4 \, \cos^2\frac{\pi}{2n} }.
$$

We can find an analytical expression
of the optimal value of $\lambda$
in the limit of large $n$.
%
In this case,
we convert Eq.
\eqref{eq:y2_sum_perfect_invt}
%and
%\eqref{eq:optimal_lambda_approx}
to an integral.
%
Particularly,
for perfect sampling,
we get from Eq. \eqref{eq:Gamma_perfect},
$$
\begin{aligned}
\Err
&=
\frac{2 \, n}{\pi \, t}
\int_0^{\pi/2}
\frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
{ \lambda \, \left(2 - 8 \, \mu_1 \, \cos^2 p - \lambda \right) }
\, dp
\notag \\
&=
\frac{n}{4 \, t}
\left(
  \frac{2 - 4 \, \mu_1 + \lambda}{ \lambda }
  +
  \frac{ \lambda }
  { \sqrt{ (2 - \lambda) (2 - 8 \, \mu_1 - \lambda) } }
\right).
\end{aligned}
$$
\note{The integral is evaluated by contour integration.
  %
$$
\begin{aligned}
\Err
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\int_0^{2 \, \pi}
\frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
{ 2 - 8 \, \mu_1 \, \cos^2 p - \lambda }
\, dp
\\
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\oint
\frac{ \left[1 - \mu_1 \, \left(z+\frac{1}{z}\right)^2 \right]^2 }
{ 2 - 2 \, \mu_1 \, \left(z + \frac{1}{z}\right)^2 - \lambda }
\, \frac{dz}{z},
\end{aligned}
$$
where the contour is the unit circle
in the complex plane of $z$.
%
This integral has five poles, one at the origin,
which produces a residue $\frac{2 - 4 \, \mu_1 - \lambda}{4}$
(from series expansion of small $z$),
and the other four at
$$
z + \frac{1}{z} = \pm\sqrt\frac{2-\lambda}{2 \, \mu_1},
$$
with two of them inside the unit circle:
$$
z_\pm = \pm \frac{\sqrt{2-\lambda} -\sqrt{2 - 8 \, \mu_1 - \lambda}}
{\sqrt{8 \mu_1}}.
$$
Thus,
$$
\begin{aligned}
\Err
&=
\frac{n}{\lambda \, t}
\left(
 \frac{2 - 4 \, \mu_1 - \lambda}{4}
 +
 \sum_{z = z_{\pm} }
 \frac{ \left(z^2 - \mu_1 (1 + z^2)^2 \right)^2 }
 { 2 z^4 (2 - 4 \mu_1 - \lambda - 4 \mu_1 z^2) }
\right)
\\
&=
\frac{n}{\lambda \, t}
\left(
  \frac{2 - 4 \, \mu_1 - \lambda}{4}
 +
 \frac{ \left(1 - \mu_1 \left(z + \frac{1}{z} \right)^2 \right)^2 }
 { 2 - 4 \mu_1 - \lambda - 4 \mu_1 z_{\pm}^2 }
\right)
\\
&=
\frac{n}{\lambda \, t}
\left(
  \frac{2 - 4 \, \mu_1 - \lambda}{4}
 +
 \frac{ (\lambda/2)^2 }
 { \sqrt{(2-\lambda) (2 - 8 \mu_1 -\lambda)} }
\right).
\end{aligned}
$$
}
%
Minimizing the function yields
%
\begin{equation}
\lambda = \frac{1 - 4 \, \mu_1} { 1 - 2 \, \mu_1 },
\label{eq:lambda_tridiag}
\end{equation}
%
and the error in the optimal is given by
%
\begin{equation}
\Err
=
\frac{n}{t}
\left(
  1+ \frac{2 \, \mu_1^2}{1-4 \, \mu_1}
\right).
\label{eq:error_tridiag}
\end{equation}
%
%From Eq. \eqref{eq:error_tridiag}, it is clear that
Clearly, the error
increases rapidly with magnitude of $\mu_1$,
no matter its sign,
and the minimal value, obtained at $\mu_1 = 0$,
corresponds to the single-bin scheme.

@@

For the Gaussian wave,
$\mu(x) \propto \exp(-\frac{1}{2} x^2/\sigma^2)$.
%
If $\Delta x \ll \sigma \ll 1$,
we can extended
the upper limit of the integrals
in Eqs. \eqref{eq:lambda_int}
and \eqref{eq:mx_normalization}
to infinity, and
%
\begin{equation}
\mu(x)
=
\frac{            1            }
     { \sqrt{ 2 \pi \sigma^2 } }
%
\exp\left(
      -\frac{       x^2     }
            { 2 \, \sigma^2 }
    \right),
\end{equation}
%
%
The eigenvalues are given by
%
\begin{equation}
\lambda_k
=
\exp\left(
      -\frac{ (k - 1)^2 \, \sigma^2 }
            {           2           }
    \right).
\label{eq:lambda_Gaussian}
\end{equation}
%
The eigenvalues are all positive,
suggesting stability of the updating scheme.

However,
as $\sigma/\Delta x \gg 1$,
the smallest eigenvalue
%
$$
\lambda_n
\approx
\exp\left(
      -\frac{ n^2 \, \sigma^2 }
            {        2        }
    \right)
=
\exp\left[
      -\frac{ \pi^2 }{ 2 }
      \left(
        \frac{  \sigma }
             { \Delta x }
      \right)^2
    \right]
.
$$
%
is exponentially small.
%
This means that
the optimal value of $\lambda$
given by Eq. \eqref{eq:optimal_lambda_approx}
would be too small in practice,
for it requires $\lambda < 2 \, \lambda_n$.
%
In other words,
with a reasonable $\lambda$,
the error of the last few fluctuation modes
will always decay suboptimally,
accordingly to Eq. \eqref{eq:y2_sum_perfect_invt}.
%
However, since these modes
represent short wavelength fluctuations,
and these modes may be in
the system under consideration.
%
Thus,
we can truncate the sum of the error function
in Eq. \eqref{eq:y2_sum_perfect_invt}
at some $k_{\max}$,
which corresponds to a minimal length scale
$l_{\min} = \Delta x \, (n /k_{\max}) = \pi/k_{\max}$
greater than $\sigma$.
%
Then,
$$
\lambda_{ k_{\max} }
\approx
\exp\left(
      -\frac{ k_{\max}^2 \, \sigma^2 }
            {           2            }
    \right)
=
\exp\left(
      -\frac{  \pi^2 \, \sigma^2    }
            {    2 \, l_{\min}^2  }
    \right).
$$

@@

\subsection{Optimal $\alpha(t)$}

@@

1. Single-bin scheme.
Show the error versus $\lambda$.

Change upon using local sampling scheme.

2. Multiple-bin scheme.
Error vs. $\lambda$.

3. Optimality of the Single-bin scheme.
Error vs. $\mu_1$.

>>>>

The Wang-Landau (WL) algorithm\cite{wang2001, wang2001pre}
and metadynamics\cite{laio2002, laio2008, barducci2011, sutto2012}
are two widely-used techniques for this purpose,
in which a bias potential is actively built-up on-the-fly
during simulation.
%
In essence, these techniques
encourage the system to leave a region that has been visited
by elevating the bias potential there.
%
A main difference of the two techniques
is that in the WL case, an update is limited to the bin
containing the current $z$,
while metadynamics adopts an extended window
covering several neighboring bins.
%
The latter ensures a smooth profile
suitable for numerical differentiation for force calculation,
and hence is more often used molecular dynamics simulations.
%
To be effective, the updates to bias potential
are made at every few steps,
while a side effect is that the underlying dynamics
is no longer a equilibrium sampling
required by statistical mechanics.
%
Thus, one has to reduce the magnitude of updating
over the course of simulation.

Naturally, the manner of reducing
the updating magnitude\cite{liang2007,
belardinelli2007, belardinelli2007jcp, belardinelli2008,
morozov2007, zhou2008, morozov2009,
komura2012, caparica2012, caparica2014,
barducci2008, dickson2011, dama2014, dickson2015}
determines the precision of the final bias potential,
hence that of PMF.
%
For the WL algorithm, it is found
that asymptotically the optimal magnitude of updating $\ln f$
should be given by the
inverse time\cite{liang2007,
belardinelli2007, belardinelli2007jcp, belardinelli2008,
morozov2007, zhou2008},
where ``time'' is defined as
the number of simulation steps
divided by the number of bins.

@@

Generally, we may wish to sample
a nearly flat distribution\cite{dayal2004, trebst2004, singh2011},
$p_i$,
by adjusting the bias potential $V_i$
on the fly.
%
For example,
we can monitor the histogram accumulated over
a period of time and use it for $\pi_i$ in
Eq. \eqref{eq:pi_p_phi1},
and then update
$V_i$ according to
$$
V_i^{\mathrm{new}}
=
V_i^{\mathrm{old}}
-
\log \frac{ p_i } { \pi_i }.
$$
%
%From Eq. \eqref{eq:pi_p_phi1},
%it is clear that to achieve the target distribution
%$\pi_i = p_i$ requires
%%
%\begin{equation}
%\log p^*_i
%=V_i + \ln p_i + c
%\label{eq:bias_pmf}
%\end{equation}
%%
%%
%In the end of the simulation,
%the PMF, $-\log p^*_i$,
%can be deduced from Eq. \eqref{eq:bias_pmf}.
%
The above inversion process,
adopted by early multicanonical or
entropic sampling\cite{berg1992, lee1993},
can be inconvenient in practice.
%
For example, a reasonably precise the histogram
requires an accumulation period
longer than the autocorrelation time,
whereas the latter is difficult to
estimate beforehand for an unknown system.

@@

In the WL algorithm\cite{wang2001, wang2001pre},
$v_i$ are updated
in each MC step $t$,
and $v_i(t)$ as a function of $t$
is updated as
%
\begin{equation}
  v_i(t+1)
  =
  v_i(t)
  +
  \delta_{i, \, i(t)}
  \frac{ \alpha(t) } { p_i }.
  \label{eq:wl_update}
\end{equation}
%
where $i(t)$ is the bin at step $t$,
and $\alpha(t)$ is the updating magnitude.
%
We refer to the above scheme as the \emph{single-bin scheme},
for it applies only to the bias potential
at the current bin, $i(t)$,
%
In a more general \emph{multiple-bin scheme}, however,
several neighboring bins are updated:
%
\begin{equation}
  v_i(t+1)
  =
  v_i(t)
  +
  w_{i, \, i(t)}
  \frac{ \alpha(t) } { p_i },
  \label{eq:mbin_update}
\end{equation}
%
where $w_{ij}$ are a set of coefficients satisfying
$\sum_{j=1}^n w_{ij} = 1$.
%
We shall discuss other restrictions on $w_{ij}$
in Sec. \ref{sec:updating-matrix}.
%
The multiple-bin scheme is more commonly
used in metadynamics.

One can show that with a constant $\alpha(t) = \alpha > 0$,
the distribution collected from
the trajectory is identical to $p_i$.
%
However, the above updates
change $v_i(t)$ continuously,
and Eq. \eqref{eq:pi_p_v} no longer holds
at all times.
%
In other words,
the error in $v_i(t)$
depends on the updating magnitude\cite{
  zhou2005, liang2005, bussi2006, liang2007,
  morozov2007, zhou2008, morozov2009, crespo2010,
  atchade2011, fort2015}.
%%
%The source of the deviation is two-fold.
%%
%First, since $v_i(t)$ is updated continuously,
%there is a random noise that is proportional
%to $\sqrt \alpha$.
%%
%Second, there is a systematic error
%that comes from the updating dynamics itself,
%as it breaks the Markovian nature
%that underlies Eq. \eqref{eq:pi_p_v}.
%%
%
Thus, one has to decrease $\alpha$ over time
to reduce the error of $v_i$,
in order to get the unbiased distribution
$\ln p_i^*$ via Eq. \eqref{eq:v_def}.

In the original WL scheme,
the updating magnitude $\alpha$ (or $\ln f$
in the original paper) is kept as a constant
for a period of time,
which is referred to as a stage below.
%
The histogram collected in the stage is monitored.
%
Once the histogram is sufficiently flat,
we are allowed to enter a new stage
with a reduced $\alpha$\cite{wang2001, wang2001pre}
(usually half as large as
that in the previous stage).
%
This scheme works well for early stages.
%
However, in later stages, it tends to reduce $\alpha$
too quickly, making the asymptotic error
saturate\cite{belardinelli2007, belardinelli2007jcp, belardinelli2008}.
%
Similarly, for metadynamics,
there are attempts to reduce the updating magnitude\cite{
  barducci2008, dickson2011, dama2014, dickson2015}.

@@

Particularly, if $\mu_2 = \dots = \mu_b = 0$,
the matrix $\mathbf w$ is tridiagonal,
%%
%\begin{equation}
%\arraycolsep=3.6pt\def\arraystretch{1.4}
%\mathbf w
%=
%\left(
%  \begin{array}{cccccccc}
%    1 - \mu_1   & \mu_1 & 0 & \dots & 0 \\
%    \mu_1 & 1 - 2 \, \mu_1  & \mu_1 & \dots & 0 \\
%    \vdots & &  & & \vdots \\
%    0 & \dots & \mu_1 & 1 - 2 \, \mu_1  & \mu_1 \\
%    0 & \dots & 0 & \mu_1 & 1 - \mu_1
%  \end{array}
%\right).
%\label{eq:wtridiag}
%\end{equation}
%%
%We have
and the eigenvalues are
\begin{equation}
  \lambda_k = 1 - 4 \, \mu_1 \sin^2 \frac{(k - 1) \, \pi}{2n}.
\label{eq:wtridiag_eigenvalue}
\end{equation}
%
Thus, for this updating matrix to be stable,
we need
$\min \lambda_k = \lambda_n > 0$,
and
$$
\mu_1 < \frac{1}{4 \, \cos^2\frac{\pi}{2n} }.
$$

We can find an analytical expression
of the optimal value of $\lambda$
in the limit of large $n$.
%
In this case,
we convert Eq.
\eqref{eq:y2_sum_perfect_invt}
%and
%\eqref{eq:optimal_lambda_approx}
to an integral,
$\sum_i \to \frac{2 \, n}{\pi} \int dp$, and
$$
\begin{aligned}
\Err_2
&=
\frac{2 \, n}{\pi \, t}
\int_0^{\pi/2}
\frac{ \Gamma(p) \, \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
{ \lambda \, \left(2 - 8 \, \mu_1 \, \cos^2 p - \lambda \right) }
\, dp.
\end{aligned}
$$

For perfect sampling,
we have Eq. \eqref{eq:Gamma_perfect}, and
$$
\begin{aligned}
\Err_2
&=
\frac{2 \, n}{\pi \, t}
\int_0^{\pi/2}
\frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
{ \lambda \, \left(2 - 8 \, \mu_1 \, \cos^2 p - \lambda \right) }
\, dp
\notag \\
&=
\frac{n}{4 \, t}
\left(
  \frac{2 - 4 \, \mu_1 + \lambda}{ \lambda }
  +
  \frac{ \lambda }
  { \sqrt{ (2 - \lambda) (2 - 8 \, \mu_1 - \lambda) } }
\right).
\end{aligned}
$$
\note{The integral is evaluated by contour integration.
  %
$$
\begin{aligned}
\Err_2
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\int_0^{2 \, \pi}
\frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
{ 2 - 8 \, \mu_1 \, \cos^2 p - \lambda }
\, dp
\\
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\oint
\frac{ \left[1 - \mu_1 \, \left(z+\frac{1}{z}\right)^2 \right]^2 }
{ 2 - 2 \, \mu_1 \, \left(z + \frac{1}{z}\right)^2 - \lambda }
\, \frac{dz}{z},
\end{aligned}
$$
where the contour is the unit circle
in the complex plane of $z$.
%
This integral has five poles, one at the origin,
which produces a residue $\frac{2 - 4 \, \mu_1 - \lambda}{4}$
(from series expansion of small $z$),
and the other four at
$$
z + \frac{1}{z} = \pm\sqrt\frac{2-\lambda}{2 \, \mu_1},
$$
with two of them inside the unit circle:
$$
z_\pm = \pm \frac{\sqrt{2-\lambda} -\sqrt{2 - 8 \, \mu_1 - \lambda}}
{\sqrt{8 \mu_1}}.
$$
Thus,
$$
\begin{aligned}
\Err_2
&=
\frac{n}{\lambda \, t}
\left(
 \frac{2 - 4 \, \mu_1 - \lambda}{4}
 +
 \sum_{z = z_{\pm} }
 \frac{ \left(z^2 - \mu_1 (1 + z^2)^2 \right)^2 }
 { 2 z^4 (2 - 4 \mu_1 - \lambda - 4 \mu_1 z^2) }
\right)
\\
&=
\frac{n}{\lambda \, t}
\left(
  \frac{2 - 4 \, \mu_1 - \lambda}{4}
 +
 \frac{ \left(1 - \mu_1 \left(z + \frac{1}{z} \right)^2 \right)^2 }
 { 2 - 4 \mu_1 - \lambda - 4 \mu_1 z_{\pm}^2 }
\right)
\\
&=
\frac{n}{\lambda \, t}
\left(
  \frac{2 - 4 \, \mu_1 - \lambda}{4}
 +
 \frac{ (\lambda/2)^2 }
 { \sqrt{(2-\lambda) (2 - 8 \mu_1 -\lambda)} }
\right).
\end{aligned}
$$
}
%
Minimizing the function yields
%
\begin{equation}
\lambda = \frac{1 - 4 \, \mu_1} { 1 - 2 \, \mu_1 },
\label{eq:lambda_tridiag_perfect}
\end{equation}
%
and the error in the optimal is given by
%
\begin{equation}
\Err
=
\frac{n}{t}
\left(
  1+ \frac{2 \, \mu_1^2}{1-4 \, \mu_1}
\right).
\label{eq:error_tridiag_prefect}
\end{equation}
%
%From Eq. \eqref{eq:error_tridiag}, it is clear that
Clearly, the error
increases rapidly with magnitude of $\mu_1$,
no matter its sign,
and the minimal value, obtained at $\mu_1 = 0$,
corresponds to the single-bin scheme.

For one-step sampling,
we have Eq. \eqref{eq:Gamma_onestep}, and
$$
\begin{aligned}
\Err_2
&=
\frac{   2 \, n }
     { \pi \, t }
\int_0^{\pi/2}
    \frac{          \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2           }
         { \lambda \, \left(2 - 8 \, \mu_1 \, \cos^2 p - \lambda \right) }
    \frac{ \cos^2 p }
         { \sin^2 p + \delta^2 }
\, dp
\notag \\
&
\stackrel{    \delta \ll 1     }
         { =\joinrel=\joinrel= }
\frac{  2 \, n  }
     { \pi \, t }
\frac{          (1 - 4 \, \mu_1)^2            }
     {  \lambda \, (2 - 8 \, \mu_1 - \lambda) }
\frac{    1   }
     { \delta }
,
\end{aligned}
$$
%
where $\delta \propto 1/n$ is a small parameter to keep the integral finite.
%
The optimal $\lambda$ is therefore
%
\begin{equation}
\lambda = 1 - 4 \, \mu_1,
\label{eq:lambda_tridiag_onestep}
\end{equation}

@@

For the Gaussian wave,
$\mu(x) \propto \exp(-\frac{1}{2} x^2/\sigma^2)$.
%
If $\Delta x \ll \sigma \ll 1$,
we can extended
the upper limit of the integrals
in Eqs. \eqref{eq:lambda_int}
and \eqref{eq:mx_normalization}
to infinity, and
%
\begin{equation}
\mu(x)
=
\frac{            1            }
     { \sqrt{ 2 \pi \sigma^2 } }
%
\exp\left(
      -\frac{       x^2     }
            { 2 \, \sigma^2 }
    \right),
\end{equation}
%
%
The eigenvalues are given by
%
\begin{align}
\lambda_k
&=
\exp\left[
      -\frac{ (k - 1)^2 \, \sigma^2 }
            {           2           }
    \right]
\notag
\\
&=
\exp\left[
      -\frac{ \pi^2 }{ 2 }
      \left(
        \frac{ k - 1 }
             {   n   }
      \right)^2
      \left(
        \frac{  \sigma }
             { \Delta x }
      \right)^2
    \right]
.
\label{eq:lambda_Gaussian}
\end{align}
%
The eigenvalues are all positive,
suggesting stability of the updating scheme.

%However,
%as $\sigma/\Delta x \gg 1$,
%the smallest eigenvalue
%%
%$$
%\lambda_n
%\approx
%\exp\left(
%      -\frac{ n^2 \, \sigma^2 }
%            {        2        }
%    \right)
%=
%\exp\left[
%      -\frac{ \pi^2 }{ 2 }
%      \left(
%        \frac{  \sigma }
%             { \Delta x }
%      \right)^2
%    \right]
%.
%$$
%%
%is exponentially small.
%%
%This means that
%the optimal value of $\lambda$
%given by Eq. \eqref{eq:optimal_lambda_approx}
%would be too small in practice,
%for it requires $\lambda < 2 \, \lambda_n$.
%%
%In other words,
%with a reasonable $\lambda$,
%the error of the last few fluctuation modes
%will always decay suboptimally,
%accordingly to Eq. \eqref{eq:y2_sum_perfect_invt}.
%%
%However, since these modes
%represent short wavelength fluctuations,
%and these modes may be in
%the system under consideration.
%%
%Thus,
%we can truncate the sum of the error function
%in Eq. \eqref{eq:y2_sum_perfect_invt}
%at some $k_{\max}$,
%which corresponds to a minimal length scale
%$l_{\min} = \Delta x \, (n /k_{\max}) = \pi/k_{\max}$
%greater than $\sigma$.
%%
%Then,
%$$
%\lambda_{ k_{\max} }
%\approx
%\exp\left(
%      -\frac{ k_{\max}^2 \, \sigma^2 }
%            {           2            }
%    \right)
%=
%\exp\left(
%      -\frac{  \pi^2 \, \sigma^2    }
%            {    2 \, l_{\min}^2  }
%    \right).
%$$

@@

\subsection{General simulation protocol}

@@

\subsection{Single-bin (WL) updating scheme}

Show the error versus $\lambda$.

Good agreement between theory and simulation.

The optimal $\lambda$ is near $1.0$,
independent of the underlying sampling scheme.

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=0.95\linewidth]{fig/errsbin.pdf}
  }
  \caption{
    \label{fig:err_singlebin}
    Error, $E$, versus the proportionality constant, $1/\lambda$,
    in Eq. \eqref{eq:alpha_trial}
    for the single-bin (WL) scheme.
    %
  }
\end{center}
\end{figure}

\subsection{Nearest neighbor updating scheme}

Normalized error vs. $\lambda$.

The optimal $\lambda$ depends on the simulation length.
This dependence is more significant in the global sampling scheme.

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=0.95\linewidth]{fig/errnnbr.pdf}
  }
  \caption{
    \label{fig:err_nnbr}
    Normalized error, $(t + t_0) \, E$,
    versus the proportionality constant, $1/\lambda$,
    in Eq. \eqref{eq:alpha_trial}
    for the nearest-neighbor updating scheme with $\mu_1 = 0.24$.
    %
  }
\end{center}
\end{figure}

\subsection{Gaussian (metadynamics) updating scheme}

\subsection{Optimality of the Single-bin scheme}

Randomize the initial error?

Error vs. $\mu_1$.

Error vs. $\sigma$.
